Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-08-03 FDT2020/main-section.tex
Questions on the Christofides part of the experiments

First we need to describe the polyhedral version Christofides' algorithm. We do this specifically for the Carr-Vempala points. Let $x$ be a Carr-Vempala points defined on a graph $G=(V,E)$. It is well known that $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull of incidence vectors of spanning trees of $G$. Hence, we can write $\frac{|V|-1}{|V|}\cdot x= \sum_{i=1}^{k}\lambda_i\chi^{T_i}$ where $T_i$ is spanning tree of $G$, $\sum_{i=1}^{k}\lambda_i=1$, and $\lambda_i\geq 0$ for $i\in [k]$. Let $O_i$ be the set of odd degree vertices of $T_i$. It is easy to see that $\frac{x}{2}$ is in the convex hull of incidence vectors of $O_i$-join of $G$, namely $O_i\join(G)$~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, and $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$.}.

 We then solve the following LP that allows us to find parity corrections that are good for the whole convex combination.
\begin{equation}\label{ojoinaverage}
\min \{ \alpha:\;\sum_{i=1}^{k} \lambda_i y^i = \alpha \cdot x,\;  y^i \in O_i\join(G) \; \mbox{for $i\in [k]$}\}.
\end{equation}
The variables in the above LP are $y^i\in \mathbb{R}^{E_x}_{
\geq 0}$ for $i\in [k]$. For each $i\in [k]$ we have $y^i\in O_i\join(G_x)$. This formulation allows the instance specific approximation ratio of Christofides' algorithm to be below $\frac{3}{2}$. Recall that a Carr-Vempala point consists of a signle cycle of fractional edges.  Figure \ref{fdtvschris} shows FDT's solutions on all Carr-Vempala points that have 10 vertices on the cycle formed by fractional edges. We show for these points the apporoximation factor provided by FDT is always better than those from the polyhedral version of Christofides' algorithm. In more details, in Figure \ref{fdtvschris} the horizontal axis of the plot is indexed with the 60 Carr-Vempala points that we considered. For each Carr-Vempala point $x$, there are two data points. The value of the first data point depicted by a circle on the vertical axis is $\frac{|V|-1}{|V|}+\alpha$  and $\alpha$ is the optimal solution to (\ref{ojoinaverage}). 
The value of the second data point depicted by a cross on the vertical axis is $C$ where $C$ is obtained from applying Theorem \ref{FDT2EC} to $x$. In other words, Figure \ref{fdtvschris} is comparing the instance-specific upper bound on integrality gap certified by Christofides' algorithm to the approximation factor of the FDT algorithm for 2EC.

>>>>

We first describe the polyhedral version of Christofides' algorithm specifically for Carr-Vempala points.
\cindy{I think it would be very helpful for the reader for you to remind them of the classic (graph-based, not polyhedral) metric (triangle-inequality) Christofides algorithm here and argue that it applies to 2EC.  Most people will remember Christofides' TSP approximation algorithm because it's so classic.  But it's easy to remind people.  Then it would be helpful to explain that when you give the LP in (\ref{ojoinaverage}) how that relates to Christofides.  If this paper is accepted, all kinds of CS/OR researchers may want to read it, so we should make their job as easy as possible.  It would be nice for others to use FDT and to build off it.}
Let $x$ be a Carr-Vempala point defined on a graph $G=(V,E)$. It is well known that $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull of incidence vectors of spanning trees of $G$.
\cindy{The statement before this depends very much on the audience.  For those who study polyhedral combinatorics especially for TSP and who know what a Carr-Vempala point is off the top of their heads, it may be well known.  For people who generally do combinatorial optimization, I would not expect them to know this. So this kind of phrasing might annoy a reader.  General rule: either a statement is really obvious (so just state it), or it is not (so give a reference or justification). Otherwise (and I know you don't intend this), you are implying the reader is stupid if they don't see the argument right away. Of course, it's the authors' job to teach and convince the reader. Thus one should avoid phrases like ``It's obvious/trivial that...'' and similar phrases. I would change the sentence before this comment to ``The vector $\frac{|V|-1}{|V|}\cdot x$ is in the convex hull...'' and give a short justification or reference. Or ``Foo and Barr (cite Foo and Bar's paper) argue that ... is in the convex hull...'' I'm sure I could figure this out if a had a little time to think, but I don't have that time now.}
Hence, we can write $\frac{|V|-1}{|V|}\cdot x= \sum_{i=1}^{k}\lambda_i\chi^{T_i}$ where $T_i$ is spanning tree of $G$, $\chi^{T_i}$ is its incidence vector, $\sum_{i=1}^{k}\lambda_i=1$, and $\lambda_i\geq 0$ for $i\in [k]$. Let $O_i$ be the set of odd-degree vertices of $T_i$. It is easy to see that $\frac{x}{2}$ is in the convex hull of incidence vectors of $O_i$-join of $G$, namely $O_i\join(G)$~\footnote{For graph $G=(V,E)$ and $O\subseteq V$ with $|O|$ even, an $O$-join of $G$ is a subgraph of $G$ that has odd degree on the vertices in $O$ and even degree on vertices in $V\setminus O$.}.
\cindy{Here is another instance of ``It's easy to see that...'' Given that (quite reasonably) you do not expect the reader to know what an O-join is, you should probably just give the short justification for this statement and remove ``It's easy to see that.'' Also, do you mean that $x/2$ is in the convex hull of the incident vectors of all $O_i$-joins for any spanning tree $T_i$?  If so, it would be good to make that clear.  Since I will not make it through every word of Sections 2-4 if you submit today, I recommend looking for similar places in those sections.}

\cindy{Do you have to precompute the joins?  Does that mean generating spanning trees before setting up the LP?  Or generating them on the fly?  Only minimum-weight spanning trees?  I haven't had a lot of time to think about this.}
 We solve the following LP that allows us to find parity corrections that are good for the whole convex combination.
\begin{equation}\label{ojoinaverage}
\min \{ \alpha:\;\sum_{i=1}^{k} \lambda_i y^i = \alpha \cdot x,\;  y^i \in O_i\join(G) \; \mbox{for $i\in [k]$}\}.
\end{equation}
The variables in the above LP are $y^i\in \mathbb{R}^{E_x}_{
\geq 0}$ for $i\in [k]$. For each $i\in [k]$ we have $y^i\in O_i\join(G_x)$. This formulation allows the instance specific approximation ratio of Christofides' algorithm to be below $\frac{3}{2}$. Recall that a Carr-Vempala point consists of a signle cycle of fractional edges.  Figure \ref{fdtvschris} shows FDT's solutions on all Carr-Vempala points that have 10 vertices on the cycle formed by fractional edges. We show for these points the apporoximation factor provided by FDT is always better than those from the polyhedral version of Christofides' algorithm. In more details, in Figure \ref{fdtvschris} the horizontal axis of the plot is indexed with the 60 Carr-Vempala points that we considered. For each Carr-Vempala point $x$, there are two data points. The value of the first data point depicted by a circle on the vertical axis is $\frac{|V|-1}{|V|}+\alpha$  and $\alpha$ is the optimal solution to (\ref{ojoinaverage}). 
The value of the second data point depicted by a cross on the vertical axis is $C$ where $C$ is obtained from applying Theorem \ref{FDT2EC} to $x$. In other words, Figure \ref{fdtvschris} is comparing the instance-specific upper bound on integrality gap certified by Christofides' algorithm to the approximation factor of the FDT algorithm for 2EC.
